---
title: "Meta-analysis investigating P and N*P addition impacts on plant functional traits"
author: "Evan Perkowski"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Script that explores the effect of P addition on leaf and whole-plant functional traits using P fertilization and N\*P fertilization experiments. The meta-analysis includes data from the MESI database as of January 07, 2025 and additional experiments compiled by Evan Perkowski.

Script summarizes the number of observations per trait and then conducts a meta-analysis to summarize plant responses to P addition, then conducts a second meta-analysis that summarizes responses when P is added in concert with N.


NOTE: NEED TO CHECK THAT ADDED SAMPLE SIZE OF ADDED EXPERIMENTS IS APPROPRIATE FOR MESI (SHOULD BE LISTED AS 'PLOT' REPS, NOT INDIVIDUAL REP NUMBER)

```{r, warning = FALSE, message=FALSE}
# Libraries
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(readr)
library(metafor)
library(MAd)

# MESI data
df_mesi <- read_csv("../data/mesi_main.csv")

# Manual data compilation
df_manual <- read_csv("../data/CNP_meta_data.csv") %>%
  mutate(sampling_year = as.character(sampling_year),
         treatment = ifelse(treatment == FALSE, "f", treatment))

# Merge MESI database with manual data compilation
df_total <- df_mesi %>%
  full_join(df_manual)

```

# Explore data availability in combined dataset for P-fertilization experiments

```{r}
explore_pfert_exps <- df_total %>%
  
  # fertilisation experiments only
  filter(treatment == "f") %>%
  
  # P-fertilisation only (without N or K addition)
  filter(npk == "_010")

head(explore_pfert_exps)

## How many experiments?
length(unique(explore_pfert_exps$exp))

## What traits are available?
unique(explore_pfert_exps$response)

```

## Select variables

```{r}
use_response_p <- c("total_biomass",
                    "agb",
                    "bgb",
                    "leaf_n_mass",
                    "leaf_n_area",
                    "leaf_p_mass",
                    "leaf_p_area",
                    "leaf_np",
                    "gs",
                    "lma",
                    "sla",
                    "spad",
                    "amax",
                    "vcmax",
                    "jmax",
                    "leaf_nue",
                    "leaf_pue",
                    "rd",
                    "tpu",
                    "asat",
                    "cica",
                    "ci"
)

pfert_responses <- explore_pfert_exps %>% 
  filter(response %in% use_response_p) %>% 
  mutate(myvar = response) %>% 
  mutate(myvar = ifelse(myvar %in% c("cica", "ci"),
                        "cica", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("asat", "amax"), 
                        "anet", myvar))

use_vars_p <- unique(pfert_responses$myvar)
```

## Analysis

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).

```{r}
pfert_responses2 <- pfert_responses %>%
  
  ## keep only essential variables and drop rows containing missing values for 
  ## essential variables
  select(id, duplicate_id, exp, myvar, treatment, sampling_year, 
         x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )

head(pfert_responses2)
```

```{r}
# Aggregate all measurements (multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
pfert_responses3 <- pfert_responses2 %>% 
  
  # suggested addition by Kevin, email 02.10.2023 10:03
  dplyr::distinct(duplicate_id, x_t, x_c, .keep_all = TRUE) |> 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, myvar, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          myvar = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, myvar, logr = es, logr_var = var) %>%
  
  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    pfert_responses2 %>%
      group_by(exp, myvar, treatment) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "myvar")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) ,
          
          # merge SLA and LMA measurements by taking inverse of logr (keep SE)
          logr = ifelse(myvar == "sla", -logr, logr),
          myvar = ifelse(myvar == "sla", "lma", myvar))

head(pfert_responses3)
```

# Meta-analysis

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.

```{r, warning=FALSE, message=FALSE}
source("../helper_fxns/analyse_meta.R")

out_p  <- purrr::map(as.list(use_vars_p), 
                   ~analyse_meta(pfert_responses3 %>% 
                                   rename(var = myvar), nam_target = .))
names(out_p) <- use_vars_p
df_box_p <- purrr::map_dfr(out_p, "df_box") |> 
  left_join(
    pfert_responses3 |> 
      group_by(myvar) |> 
      summarise(logr_min = min(logr), logr_max = max(logr)) |> 
      rename(var = myvar),
    by = "var"
  )
saveRDS(df_box_p, file = paste0(here::here(), "df_box_pfert.rds"))
```

## Final data size

Number of data points (plot-level measurements) per variable:

```{r, message=FALSE}
pfert_responses3 %>% 
  group_by(myvar) %>% 
  summarise(n_plots = sum(n_c, na.rm = TRUE), n_exp = n()) %>% 
  rename("Variable"="myvar", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

Number of data points (plot-level measurements) per experiment:
```{r, warning=FALSE}
pfert_responses3 %>% 
  group_by(exp) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Experiment"="exp", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

## Some quick plots:
```{r}

pfert_responses3 <- pfert_responses3 %>%
  mutate(myvar = factor(myvar, levels = c("bgb", "agb", "total_biomass", "leaf_pue",
                                          "leaf_nue", "rd", "tpu", "jmax", "vcmax",
                                          "cica", "gs", "anet", "spad", "leaf_np",
                                          "leaf_p_area", "leaf_p_mass", "leaf_n_area",
                                           "leaf_n_mass", "lma")))

meta_plot <- ggplot(data = subset(pfert_responses3, myvar != "cica"), 
                    aes(x = myvar, y = logr)) +
  geom_jitter(color = rgb(0,0,0,0.3), 
              aes( size = 1/logr_se ), 
              position = position_jitter(w = 0.2, h = 0),
              show.legend = FALSE) +
  geom_crossbar( data = df_box_p %>% drop_na(var), 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax),
                 fill = "royalblue", 
                 color = "royalblue4", 
                 alpha = 0.6, 
                 width = 0.5 ) +
  geom_hline( yintercept = 0.0, linewidth = 0.5, linetype = "dotted" ) +
  scale_x_discrete(labels = c("Belowground biomass", 
                              "Aboveground biomass", 
                              "Total biomass", 
                              "PPUE",
                              "PNUE", 
                              expression("R"["d"]), 
                              "TPU", 
                              expression("J"["max"]), 
                              expression("V"["cmax"]),
                              expression("g"["s"]),
                              expression("A"["net"]), 
                              "SPAD", 
                              "Leaf N:P",
                              expression("P"["area"]), 
                              expression("P"["mass"]), 
                              expression("N"["area"]),
                              expression("N"["mass"]), 
                              expression("M"["area"]))) +
  
  #scale_x_discrete("", labels = mylabl) +
  labs(x = "", 
       y = "Log response to P addition") +
  coord_flip() +
  theme_classic(base_size = 18)
meta_plot

```



# Explore data availability in combined dataset for N*P-fertilization experiments

```{r}
explore_npfert_exps <- df_total %>%
  
  # fertilisation experiments only
  filter(treatment == "f") %>%
  
  # P-fertilization in concert with N-fertilization (without K addition)
  filter(npk == "_110")

head(explore_npfert_exps)

## How many experiments?
length(unique(explore_npfert_exps$exp))

## What traits are available?
unique(explore_npfert_exps$response)

```

## Select variables

```{r}
use_response_np <- c("total_biomass",
                    "agb",
                    "bgb",
                    "leaf_n_mass",
                    "leaf_n_area",
                    "leaf_p_mass",
                    "leaf_p_area",
                    "leaf_np",
                    "gs",
                    "lma",
                    "sla",
                    "spad",
                    "amax",
                    "vcmax",
                    "jmax",
                    "leaf_nue",
                    "leaf_pue",
                    "rd",
                    "tpu",
                    "asat",
                    "cica",
                    "ci"
)

npfert_responses <- explore_npfert_exps %>% 
  filter(response %in% use_response_np) %>% 
  mutate(myvar = response) %>% 
  mutate(myvar = ifelse(myvar %in% c("cica", "ci"),
                        "cica", myvar)) %>% 
  mutate(myvar = ifelse(myvar %in%  c("asat", "amax"), 
                        "anet", myvar))

use_vars_np <- unique(npfert_responses$myvar)
```

## Analysis

Calculate `"ROM"` - the log transformed ratio of means (Hedges et al., 1999; Lajeunesse, 2011) for each observation pair (ambient and elevated).

```{r}
npfert_responses2 <- npfert_responses %>%
  
  ## keep only essential variables and drop rows containing missing values for 
  ## essential variables
  select(id, duplicate_id, exp, myvar, treatment, sampling_year, 
         x_t, x_c, sd_t, sd_c, rep_t, rep_c) %>%
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = x_t, sd1i = sd_t, n1i = rep_t, 
    m2i = x_c, sd2i = sd_c, n2i = rep_c, 
    data = ., 
    append = TRUE, 
    var.names = c("logr", "logr_var") 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  ## get standard error
  mutate( logr_se = sqrt(logr_var) / sqrt(rep_t) )

head(npfert_responses2)
```


```{r}
# Aggregate all measurements (multiple years, sampling dates and plots) by experiment (and response variable - although here only one) for meta analysis.
npfert_responses3 <- npfert_responses2 %>% 
  
  # suggested addition by Kevin, email 02.10.2023 10:03
  dplyr::distinct(duplicate_id, x_t, x_c, .keep_all = TRUE) |> 
  
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  
  # re-create ID (common ID per experiment and response variable)
  select(-id) %>%
  mutate( id = paste(exp, myvar, sep = "_XXX_")) %>% 
  
  MAd::agg( 
    id = id, 
    es = logr, 
    var = logr_var,
    cor = 1.0, 
    method = "BHHR", 
    data = . 
  ) %>% 
  
  ## to keep the output readable from the console output
  as_tibble() %>% 
  
  # separate ID again for ease of data use
  mutate( id = str_split(id, "_XXX_") ) %>%
  mutate( exp = purrr::map_chr(id, 1),
          myvar = purrr::map_chr(id, 2) ) %>%
  
  ## rename again
  select(exp, myvar, logr = es, logr_var = var) %>%
  
  ## add number of observations (sum of plots and repeated samplings)
  left_join(
    npfert_responses2 %>%
      group_by(exp, myvar, treatment) %>%
      summarise(n_c = sum(rep_c), n_t = sum(rep_t)),
    by = c("exp", "myvar")
  ) %>% 
  
  ## get standard error. Verify if number available observations are identical
  ## for ambient and elevated. Use N from control here (n_c).
  mutate( logr_se = sqrt(logr_var) / sqrt(n_c) ,
          
          # merge SLA and LMA measurements by taking inverse of logr (keep SE)
          logr = ifelse(myvar == "sla", -logr, logr),
          myvar = ifelse(myvar == "sla", "lma", myvar))

head(npfert_responses3)
```

# Meta-analysis

Aggregate log-ratios across multiple experiments, taking into account their respective variance and using the experiment identity as a grouping factor for random intercepts.

```{r, warning=FALSE, message=FALSE}
out_np  <- purrr::map(as.list(use_vars_np), 
                   ~analyse_meta(npfert_responses3 %>% 
                                   rename(var = myvar), nam_target = .))
names(out_np) <- use_vars_np
df_box_np <- purrr::map_dfr(out_np, "df_box") |> 
  left_join(
    npfert_responses3 |> 
      group_by(myvar) |> 
      summarise(logr_min = min(logr), logr_max = max(logr)) |> 
      rename(var = myvar),
    by = "var"
  )
saveRDS(df_box_np, file = paste0(here::here(), "df_box_npfert.rds"))
```

## Final data size

Number of data points (plot-level measurements) per variable:

```{r, message=FALSE}
npfert_responses3 %>% 
  group_by(myvar) %>% 
  summarise(n_plots = sum(n_c, na.rm = TRUE), n_exp = n()) %>% 
  rename("Variable"="myvar", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

Number of data points (plot-level measurements) per experiment:
```{r, warning=FALSE}
npfert_responses3 %>% 
  group_by(exp) %>% 
  summarise(n_plots = sum(n_c), n_exp = n()) %>% 
  rename_("Experiment"="exp", "N observations"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

## Some quick plots:
```{r}
npfert_responses3 <- npfert_responses3 %>%
  mutate(myvar = factor(myvar, levels = c("bgb", "agb", "total_biomass", "leaf_pue",
                                          "leaf_nue", "rd", "tpu", "jmax", "vcmax",
                                          "cica", "gs", "anet", "spad", "leaf_np",
                                          "leaf_p_area", "leaf_p_mass", "leaf_n_area",
                                           "leaf_n_mass", "lma")))

meta_plot_np <- ggplot(data = subset(npfert_responses3, myvar != "cica"), 
                    aes(x = myvar, y = logr)) +
  geom_jitter(color = rgb(0,0,0,0.3), 
              aes( size = 1/logr_se ), 
              position = position_jitter(w = 0.2, h = 0),
              show.legend = FALSE) +
  geom_crossbar( data = df_box_np %>% drop_na(var), 
                 aes(x = var, y = middle, ymin = ymin, ymax = ymax),
                 fill = "royalblue", 
                 color = "royalblue4", 
                 alpha = 0.6, 
                 width = 0.5 ) +
  geom_hline( yintercept = 0.0, linewidth = 0.5, linetype = "dotted" ) +
  scale_x_discrete(labels = c("Belowground biomass", 
                              "Aboveground biomass", 
                              "Total biomass", 
                              "PPUE",
                              "PNUE", 
                              expression("R"["d"]), 
                              "TPU", 
                              expression("J"["max"]), 
                              expression("V"["cmax"]),
                              expression("g"["s"]), 
                              "SPAD", 
                              "Leaf N:P",
                              expression("P"["area"]), 
                              expression("P"["mass"]), 
                              expression("N"["area"]),
                              expression("N"["mass"]), 
                              expression("M"["area"]))) +
  
  #scale_x_discrete("", labels = mylabl) +
  labs(x = "", 
       y = "Log response to N*P addition") +
  coord_flip() +
  theme_classic(base_size = 18)
meta_plot_np

```



